---
title: "Using the databricks community cloud"
output: html_notebook
---

If you choose to run in a cloud environment (databricks community cloud) sparkR needs to be installed. However `ìnstall.packages("SparkR")` will fail, as the **SparkR** package is not uploaded to CRAN. You need to follow the installation instructions here: https://github.com/apache/spark/tree/master/R for manual installation. You can follow https://issues.apache.org/jira/browse/SPARK-15799 to find out if SparkR is available now from CRAN.

See https://databricks.com/blog/2017/05/25/using-sparklyr-databricks.html for general instructions of using sparkR/sparklyR with databricks.

> Note, the instructions to install sparkR might be incomplete.

To install SparkR make sure to have these R packages installed
```{r, eval=FALSE, include=TRUE}
install.packages('testthat', repos=c('https://cloud.r-project.org'))
install.packages('roxygen2', repos=c('https://cloud.r-project.org'))
```
Then build spark including the SparkR package from source. The following code snippet assumes that `git` and `maven` are installed an on the `$PATH`
```{bash, eval=FALSE, include=TRUE}
git clone https://github.com/apache/spark.git
cd spark
git checkout v2.2.0
export MAVEN_OPTS="-Xmx2g -XX:ReservedCodeCacheSize=512m"
./dev/make-distribution.sh --name my-spark --tgz -Psparkr -Phadoop-2.7 -Phive -Phive-thriftserver -Pmesos -Pyarn -Pnetlib-lgpl
```
install the freshly built R package:
```{bash, eval=FALSE, include=TRUE}
# where /home/username/R is where R is installed and /home/username/R/bin contains the files R and RScript
export R_HOME=/home/username/R
./install-dev.sh
```
and use it via:
```{bash, eval=FALSE, include=TRUE}
# Set this to where Spark is installed
Sys.setenv(SPARK_HOME="/Users/username/spark")
# This line loads SparkR from the installed directory
.libPaths(c(file.path(Sys.getenv("SPARK_HOME"), "R", "lib"), .libPaths()))
library(SparkR)
sparkR.session()
```

Finally, it should be possible to use sparklyR with databricks like:
```{r, eval=FALSE, include=TRUE}
# cloud spark https://databricks.com/blog/2017/05/25/using-sparklyr-databricks.html
# spark <- spark_connect(method = "databricks") # that also requires sparkR to be installed
```